Cooperación y ergodicidad

### 2021-08-03 article.tex
Varios: Citas, código, presentación.

\begin{abstract}
Public goods games (PGGs) are paradigmatic among such non-decomposable multiplayer games.
PGGs are models of situations where individuals face the dilemma of providing and/or maintaining a public good: a common resource that is both non-excludable (no individual can be excluded from its consumption) and non-rivalrous (one individual’s use of the public good does not diminish its availability to another individual).

Are typical examples of public goods whose abusive exploitation by non-contributing individuals may lead to the so-called tragedy of the commons, a situation in which nobody contributes and therefore no public good is produced or maintained.

By far, the most well known PGG is the N-person prisoner’s dilemma (NPD)

El juego de los bienes comunes, bajo regímenes multiplicativos comos los de la vida (secuencua de probabilidades de supervivencia y reproducción), no son dilemas del prisionero como se creía.
Incluso cuando la cooperación tiene un costo, la emergencia de estrategias parasitarias no puede invadir  hay una ventaja evolutiva 
\end{abstract}

@@

In a sense, repetition, reputation and punishment do not solve the problem of the evolution of cooperation under the prisoner’s dilemma: they simply transform the underlying game into a stag hunt (Skyrms, 2004; Taylor & Nowak, 2007). Cooperation under the stag hunt is however easier to achieve, as the problem of the stability of cooperation is now solved. The only problem now is the origin or initial viability of cooperation: given that cooperation is stable, how can it get a foothold in a population entirely comprised of defectors?

>>>>

\begin{abstract}
\end{abstract}

@@

In a sense, repetition, reputation and punishment do not solve the problem of the evolution of cooperation under the prisoner’s dilemma: they simply transform the underlying game into a stag hunt (Skyrms, 2004; Taylor \& Nowak, 2007). Cooperation under the stag hunt is however easier to achieve, as the problem of the stability of cooperation is now solved. The only problem now is the origin or initial viability of cooperation: given that cooperation is stable, how can it get a foothold in a population entirely comprised of defectors?

\section{Cites}

Mynard Smith sobre las aplicaciones de la teoría de juegos.
\begin{quotation}
    Game theory is more readily applied to biology than to the field of economic behaviour for which it was originally designed. (...).
    In seeking the solution of a game, the concept of human rationality is replaced by that of evolutionary stability.
    The advantage here is that there are good theoretical reasons to expect populations to evolve to stable states, whereas there are grounds for doubting whether human beings always behave rationally. \cite{maynardSmith1982-evolutionTheoryGames}

    A central assumption of classical game theory is that the players will behave rationally, and according to some criterion of self-interest.
    Such an assumption would clearly be out of place in an evolutionary context.
    Instead, the criterion of rationality is replaced by that of population dynamics and stability, and the criterion of self-interest by Darwinian fitness. (...).
    They lead to a new type of 'solution' to a game, the 'evolutionarily stable strategy' or ESS.
\end{quotation}

\begin{quotation}
    The central idea I am developing here is that an essential part of evolution deals with adaptation to changing environmental conditions~\cite{pigliucci2001-plasticity}.
\end{quotation}

Parafraseando a Den Boer: the chance of survival of a population depends on a reciprocity system for spreading of the risk within populations and between populations of different species.
En sus palabras
\begin{quotation}
    Which are the factors governing the distribution of a population over a number of localities?
    The daily contact with data from the field soon showed that change and variation predominate in the natural situation: populations do not fit
    nicely to some definite habitat definable in terms of important ecological factors.
    Heterogeneity and instability must not be considered as just a drawback of field data to be neglected.
    On the contrary, heterogeneity and/or instability must be recognized as fundamental features of a natural situation.
    The chance of survival of a population may even be increased, because the variation within the population makes it possible to cope with the variation in space and time of the habitat.
    This possibility led me to formulate the concept ``spreading of risk''~\cite{denBoer1968-spreadingRisk}: 1. Spreading of risk by phenotypic variation; 2. Spreading of risk in time (e.g. age classes); 3. Spreading of risk in space (e.g. different microenvironments); 4. Spreading of risk and relations to other species (monophagous vs polyphagous).
    The simultaneous occurrence of different kinds of heterogeneity can be expected to increase the spreading of risk.
    Biocoenosis rich in species may be expected to be more stable (provided with more intricate spreading of the risk) than one with few species.
    Hence, the ``balance of nature'' is not a balance but rather a relatively high degree of stability.

    It is supposed that the stabilization of animal numbers in natural populations is established by spreading of the risk, incidentally supplem ented by some kind of density limitation.
    The ``balance of nature'' must be the result of a complex and highly intricate system of spreading of the risk within populations and between populations of the same and especially of different species.
\end{quotation}

Ya Den Boer plantea como hipótesis un trade-off del ``spreading of the risk'', respecto de la densidad poblacional.
Seger se pregunta
\begin{quotation}
    Is there am evolutionary trade-off between \emph{expected} fitness and the \emph{variance} of fitness?
    Here, we show why there is only under certain circumstances~\cite{seger1987-betHedging}.

    Dampster~\cite{dempster1955-geometricMean} showed that for temporal fluctations on a scale of generations, the relevant measure of fitness is its geometric mean.
    Dampster result is the starting-point for all subsequent models of genetic evolution in temporally varying environments.

    The geometric meman is the natural measure of long-term fitness under temporal variation beacause, like population growth itself, it is inherently multiplicative rather than additive.
    It is therefore very sensitive to occasional small values.
    If there is any variation, the geometric mean will be less than the arithmetic mean.

    Although the genotype with the highest geometric-mean fitness is favoured (and certainly goes to fixation), the \emph{populations's} geometric-mean fitness is not necessarily maximized at equilibrium [same example as \cite{starrfelt2012-bet}].
\end{quotation}
Y muestra el mismo ejemplo que Kokko.

\\

Jorge Peña
\begin{quotation}
    Public goods games (PGGs) are paradigmatic among such non-decomposable multiplayer games.
    PGGs are models of situations where individuals face the dilemma of providing and/or maintaining a public good: a common resource that is both non-excludable (no individual can be excluded from its consumption) and non-rivalrous (one individual’s use of the public good does not diminish its availability to another individual).
    Are typical examples of public goods whose abusive exploitation by non-contributing individuals may lead to the so-called tragedy of the commons, a situation in which nobody contributes and therefore no public good is produced or maintained.
    By far, the most well known PGG is the N-person prisoner’s dilemma (NPD). \cite{pena2012-phd}
\end{quotation}
